---
title: "EDA Collusion 6"
author: "Ross May"
date: "2026-01-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(123)
```

## Flag homogeneous markets

```{r flag-anomalous-markets}
# Flag markets where normalized total consumer loss exceeds upper bounds
library(dplyr)
library(tidyr)
library(readr)

# Define the configurations for each case
cases <- data.frame(
  region = c("gb", "gb", "se", "se"),
  product = c("freezers", "headphones", "freezers", "headphones"),
  cl_folder = c("cl_results_gb_freezers", "cl_results_gb_headphones", 
                "cl_results_se_freezers", "cl_results_se_headphones"),
  upper_folder = c("upper_pi_bounds_gb_freezers", "upper_pi_bounds_gb_headphones",
                   "upper_pi_bounds_se_freezers", "upper_pi_bounds_se_headphones"),
  upper_file = c("tcl_upper_gb_freez.csv", "tcl_upper_gb_hp.csv",
                 "tcl_upper_se_freez.csv", "tcl_upper_se_hp.csv"),
  stringsAsFactors = FALSE
)

# Function to process a single market file (a_markets or na_markets)
process_market_file <- function(file_path, upper_bounds, region, product, market_type) {
  # Read the market data
  market_data <- read_csv(file_path, show_col_types = FALSE)
  
  # Check if file is empty or has no data rows
  if (nrow(market_data) == 0) {
    return(data.frame())
  }
  
  # Normalize tot_consum_loss by mean of tot_consum_loss
  market_data <- market_data %>%
    mutate(
      normalized_tcl = tot_consum_loss / mean(tot_consum_loss, na.rm = TRUE),
      region = region,
      product = product,
      market_type = market_type
    )
  
  # Merge with upper bounds
  market_data <- market_data %>%
    left_join(upper_bounds, by = c("mark_size" = "ms"))
  
  # Flag markets where normalized TCL exceeds upper bound
  market_data <- market_data %>%
    mutate(
      flagged = normalized_tcl > tcl_upper,
      excess_amount = normalized_tcl - tcl_upper,
      excess_pct = ((normalized_tcl - tcl_upper) / tcl_upper) * 100
    )
  
  return(market_data)
}

# Initialize list to store all results
all_results <- list()

# Process each case
for (i in 1:nrow(cases)) {
  region <- cases$region[i]
  product <- cases$product[i]
  cl_folder <- cases$cl_folder[i]
  upper_folder <- cases$upper_folder[i]
  upper_file <- cases$upper_file[i]
  
  cat(sprintf("\n=== Processing: %s %s ===\n", toupper(region), product))
  
  # Read upper bounds
  upper_bounds_path <- file.path("output", upper_folder, upper_file)
  upper_bounds <- read_csv(upper_bounds_path, show_col_types = FALSE)
  
  # Process a_markets
  a_markets_path <- file.path("output", cl_folder, "a_markets.csv")
  if (file.exists(a_markets_path)) {
    a_results <- process_market_file(a_markets_path, upper_bounds, region, product, "a_markets")
    
    if (nrow(a_results) > 0) {
      # Save detailed results for a_markets
      output_file <- file.path("output", cl_folder, "a_markets_flagged.csv")
      write_csv(a_results, output_file)
      cat(sprintf("  A-markets: %d markets, %d flagged (%.1f%%)\n", 
                  nrow(a_results), sum(a_results$flagged), 
                  100 * sum(a_results$flagged) / nrow(a_results)))
      
      all_results[[paste0(region, "_", product, "_a")]] <- a_results
    }
  }
  
  # Process na_markets
  na_markets_path <- file.path("output", cl_folder, "na_markets.csv")
  if (file.exists(na_markets_path)) {
    na_results <- process_market_file(na_markets_path, upper_bounds, region, product, "na_markets")
    
    if (nrow(na_results) > 0) {
      # Save detailed results for na_markets
      output_file <- file.path("output", cl_folder, "na_markets_flagged.csv")
      write_csv(na_results, output_file)
      cat(sprintf("  NA-markets: %d markets, %d flagged (%.1f%%)\n", 
                  nrow(na_results), sum(na_results$flagged), 
                  100 * sum(na_results$flagged) / nrow(na_results)))
      
      all_results[[paste0(region, "_", product, "_na")]] <- na_results
    }
  }
}

# Combine all results
combined_results <- bind_rows(all_results)

# Save combined detailed results
write_csv(combined_results, "output/all_markets_flagged_detailed.csv")
cat(sprintf("\n=== Overall Results ===\n"))
cat(sprintf("Total markets analyzed: %d\n", nrow(combined_results)))
cat(sprintf("Total flagged markets: %d (%.1f%%)\n", 
            sum(combined_results$flagged), 
            100 * sum(combined_results$flagged) / nrow(combined_results)))

# Create summary statistics
summary_stats <- combined_results %>%
  group_by(region, product, market_type) %>%
  summarise(
    total_markets = n(),
    flagged_markets = sum(flagged),
    flagged_pct = 100 * sum(flagged) / n(),
    avg_normalized_tcl = mean(normalized_tcl),
    avg_tcl_upper = mean(tcl_upper, na.rm = TRUE),
    max_excess_amount = max(excess_amount[flagged], na.rm = TRUE),
    avg_excess_amount = mean(excess_amount[flagged], na.rm = TRUE),
    avg_excess_pct = mean(excess_pct[flagged], na.rm = TRUE),
    .groups = "drop"
  )

# Save summary statistics
write_csv(summary_stats, "output/flagged_markets_summary.csv")
write_csv(summary_stats, "assets/flagged_markets_summary.csv")
print(summary_stats)

# Create a list of flagged markets for further analysis
flagged_markets <- combined_results %>%
  filter(flagged) %>%
  select(region, product, market_type, product_id, mark_size, 
         tot_consum_loss, normalized_tcl, tcl_upper, 
         excess_amount, excess_pct) %>%
  arrange(region, product, market_type, desc(excess_pct))

# Save flagged markets list
write_csv(flagged_markets, "output/flagged_markets_for_analysis.csv")
write_csv(flagged_markets, "assets/flagged_markets_for_analysis.csv")
cat(sprintf("\nFlagged markets saved to: output/flagged_markets_for_analysis.csv\n"))
cat(sprintf("Number of flagged markets for pricing pattern analysis: %d\n", nrow(flagged_markets)))

# Create market size distribution of flagged markets
ms_distribution <- flagged_markets %>%
  group_by(region, product, market_type, mark_size) %>%
  summarise(
    n_flagged = n(),
    avg_excess_pct = mean(excess_pct),
    .groups = "drop"
  ) %>%
  arrange(region, product, market_type, mark_size)

write_csv(ms_distribution, "output/flagged_markets_by_market_size.csv")
write_csv(ms_distribution, "assets/flagged_markets_by_market_size.csv")

# Create visual summary table
visual_summary <- combined_results %>%
  group_by(region, product) %>%
  summarise(
    a_markets_total = sum(market_type == "a_markets"),
    a_markets_flagged = sum(market_type == "a_markets" & flagged),
    a_markets_flagged_pct = 100 * sum(market_type == "a_markets" & flagged) / sum(market_type == "a_markets"),
    na_markets_total = sum(market_type == "na_markets"),
    na_markets_flagged = sum(market_type == "na_markets" & flagged),
    na_markets_flagged_pct = 100 * sum(market_type == "na_markets" & flagged) / sum(market_type == "na_markets"),
    .groups = "drop"
  )

write_csv(visual_summary, "output/flagged_markets_visual_summary.csv")
write_csv(visual_summary, "assets/flagged_markets_visual_summary.csv")
print(visual_summary)

cat("\n=== Analysis Complete ===\n")
cat("Output files created:\n")
cat("  1. output/all_markets_flagged_detailed.csv - All markets with flagging status\n")
cat("  2. output/flagged_markets_summary.csv - Summary statistics by region/product/type\n")
cat("  3. output/flagged_markets_for_analysis.csv - List of flagged markets for pricing analysis\n")
cat("  4. output/flagged_markets_by_market_size.csv - Distribution by market size\n")
cat("  5. output/flagged_markets_visual_summary.csv - Quick visual summary\n")
cat("  6. Individual *_flagged.csv files in each cl_results_* folder\n")
```

## Results Summary

The analysis identified **8 flagged markets (4.2%)** out of 192 total markets where normalized Competitive Price Gap (CPG) exceeds the upper prediction interval bounds (see Table 1 and Figure 1):

**Table 1. Summary Statistics by Region, Product, and Market Type** - Detailed breakdown showing total markets, flagged markets, flagged percentages, and excess measures for each region-product-market type combination.

**Breakdown by Region and Product:**
- **GB Freezers**: 0 flagged markets (0% of 17 total)
  - 0 in A markets (0% of 2)
  - 0 in NA markets (0% of 15)
- **GB Headphones**: 3 flagged markets (5.1% of 59 total)
  - 2 in A markets (5.4% of 37)
  - 1 in NA markets (4.5% of 22)
- **SE Freezers**: 1 flagged market (12.5% of 8 total)
  - 0 in A markets (0% of 0)
  - 1 in NA markets (12.5% of 8)
- **SE Headphones**: 4 flagged markets (3.7% of 108 total)
  - 2 in A markets (5.0% of 40)
  - 2 in NA markets (2.9% of 68)

These flagged markets exhibit anomalously high CPG relative to their upper prediction interval bounds and are candidates for further pricing pattern analysis to identify potential coordination mechanisms.

**Figure 1. Percentage of Flagged Markets by Country, Product Category, and Market Type** - Bar chart comparing flagged percentages across all region-product-market type combinations.

**Figure 2. Excess Percentage Above CPG Upper Bound for Flagged Markets** - Scatter plot showing how far each of the 8 flagged markets exceeds its respective upper bound threshold.

**Figure 3. Normalized CPG vs Upper Bound** - Faceted scatter plots for each region-product combination showing the relationship between normalized CPG and upper bounds, with flagged markets (above the diagonal line) highlighted in red.

**Table 2. Visual Summary by Region and Product** - High-level overview showing A market and NA market totals and flagged counts for each region-product combination.

**Table 3. Top 10 Flagged Markets** - Individual market details for the most extreme flagged markets, ranked by excess percentage.

**Figure 4. Proportion of Flagged Markets by Market Size** - Stacked bar charts showing the distribution of flagged vs non-flagged markets across different market sizes for each region-product combination.

**Note on Terminology:** The analysis uses "Competitive Price Gap" (CPG) as the metric, which is represented in the code as `tot_consum_loss`. This terminology reflects the gap between observed and competitive pricing that results in consumer welfare loss. Markets are classified as A markets (assortative matching) or NA markets (non-assortative matching).

## Methodology: Summary Statistics Computations

### 1. Data Processing Pipeline
The analysis processes four region-product combinations (GB/SE × Freezers/Headphones), analyzing both A markets (markets with assortative matching) and NA markets (markets with non-assortative matching) separately. Results are presented in **Table 1** (detailed statistics showing all 8 region-product-type combinations) and **Table 2** (visual summary aggregated by region-product only).

### 2. Normalization
For each market within a region-product-type combination (e.g., all GB headphones A markets), the Competitive Price Gap (CPG) is normalized by dividing by the mean:

$$\text{normalized\_tcl} = \frac{\text{tot\_consum\_loss}}{\text{mean(tot\_consum\_loss)}}$$

This produces a dimensionless measure where values around 1.0 represent markets with average CPG, while values >1 indicate above-average consumer welfare loss. The normalization occurs within each region-product-type group, allowing comparison across different market sizes within the same category.

### 3. Flagging Criterion
Markets are flagged when their normalized CPG exceeds the upper prediction interval bound for their market size:

$$\text{flagged} = \begin{cases} 
\text{TRUE} & \text{if normalized\_tcl} > \text{tcl\_upper} \\
\text{FALSE} & \text{otherwise}
\end{cases}$$

The upper bounds (`tcl_upper`) are market-size specific thresholds derived from prediction intervals, stored in files like `tcl_upper_gb_freez.csv`, and merged based on the `mark_size` variable.

### 4. Excess Measures
For flagged markets, we compute two excess measures to quantify the degree of anomaly:

- **Excess Amount**: $\text{excess\_amount} = \text{normalized\_tcl} - \text{tcl\_upper}$
- **Excess Percentage**: $\text{excess\_pct} = \frac{\text{normalized\_tcl} - \text{tcl\_upper}}{\text{tcl\_upper}} \times 100$

These measures indicate how far above the upper bound each flagged market falls, both in absolute and relative terms.

### 5. Summary Statistics by Region-Product-Type
The `summary_stats` table groups markets by region, product, and market_type, computing:

- **total_markets**: Count of markets in the group: $n()$
- **flagged_markets**: Count of flagged markets: $\sum(\text{flagged})$
- **flagged_pct**: Percentage flagged: $\frac{\sum(\text{flagged})}{n()} \times 100$
- **avg_normalized_tcl**: Mean normalized CPG: $\text{mean(normalized\_tcl)}$ (equals 1.0 by construction)
- **avg_tcl_upper**: Mean upper bound: $\text{mean(tcl\_upper)}$
- **max_excess_amount**: Maximum excess among flagged markets: $\max(\text{excess\_amount}|\text{flagged})$
- **avg_excess_amount**: Mean excess among flagged markets: $\text{mean(excess\_amount}|\text{flagged})$
- **avg_excess_pct**: Mean excess % among flagged markets: $\text{mean(excess\_pct}|\text{flagged})$

This table is exported to `assets/table1_summary_statistics.csv` with formatted column names for Word document inclusion (e.g., "Avg Normalised CPG", "CPG Upper"). **Figure 1** visualizes the flagged percentages from this table as a grouped bar chart.

### 6. Visual Summary by Region-Product
The `visual_summary` table groups by region and product only, providing a higher-level overview across market types:

- **a_markets_total**: Count of a_markets: $\sum(\text{market\_type} == \text{"a\_markets"})$
- **a_markets_flagged**: Count of flagged a_markets: $\sum(\text{market\_type} == \text{"a\_markets"} \land \text{flagged})$
- **a_markets_flagged_pct**: Percentage of a_markets flagged: $\frac{\sum(\text{market\_type} == \text{"a\_markets"} \land \text{flagged})}{\sum(\text{market\_type} == \text{"a\_markets"})} \times 100$
- **na_markets_total**: Count of na_markets: $\sum(\text{market\_type} == \text{"na\_markets"})$
- **na_markets_flagged**: Count of flagged na_markets: $\sum(\text{market\_type} == \text{"na\_markets"} \land \text{flagged})$
- **na_markets_flagged_pct**: Percentage of na_markets flagged: $\frac{\sum(\text{market\_type} == \text{"na\_markets"} \land \text{flagged})}{\sum(\text{market\_type} == \text{"na\_markets"})} \times 100$

This table is exported to `assets/table2_visual_summary.csv` with formatted column names (e.g., "A Markets Total", "NA Markets Flagged %"). This provides a high-level overview complementing the detailed breakdown in Table 1.

### 7. Market Size Distribution
The `ms_distribution` table groups flagged markets by region, product, market_type, and mark_size, showing:

- **n_flagged**: Count of flagged markets at each market size
- **avg_excess_pct**: Average excess percentage at each market size

This table is exported to `assets/flagged_markets_by_market_size.csv`. **Figure 4** visualizes the proportion of flagged markets by market size across all region-product combinations as stacked bar charts.

### 8. Exported Data Files
The `run_analysis_export.R` script generates multiple output files in the `assets/` folder:

**Tables for Word Document:**
- `table1_summary_statistics.csv` - **Table 1:** Summary Statistics by Region, Product, and Market Type (formatted with proper capitalization)
- `table2_visual_summary.csv` - **Table 2:** Visual Summary by Region and Product (formatted with proper capitalization)
- `table3_top_flagged_markets.csv` - **Table 3:** Top 10 Flagged Markets ranked by excess percentage (formatted with proper capitalization)

**Data Files:**
- `flagged_markets_summary.csv` - Raw summary statistics
- `flagged_markets_visual_summary.csv` - Raw visual summary
- `flagged_markets_for_analysis.csv` - Complete list of all flagged markets with details
- `flagged_markets_by_market_size.csv` - Distribution by market size
- `all_markets_combined_results.csv` - Full dataset with all markets and flagging status

**Figures:**
- `figure1_flagged_pct_by_region_product.png` - **Figure 1:** Percentage of Flagged Markets by Country, Product Category, and Market Type. Grouped bar chart comparing A markets vs NA markets across all region-product combinations.
- `figure2_excess_pct_flagged_markets.png` - **Figure 2:** Excess Percentage Above CPG Upper Bound for Flagged Markets. Scatter plot showing the magnitude of threshold exceedance for each of the 8 flagged markets.
- `figure3_normalized_tcl_vs_upper_bound.png` - **Figure 3:** Normalized CPG vs Upper Bound. Faceted scatter plots (one panel per region-product combination) showing all markets, with flagged markets (above the diagonal) highlighted.
- `figure4_proportion_flagged_by_market_size.png` - **Figure 4:** Proportion of Flagged Markets by Market Size. Stacked bar charts showing the relative frequency of flagged vs non-flagged markets at each market size.

**Summary Report:**
- `analysis_summary_report.txt` - Text summary of findings and methodology

### Interpreting NaN and -Inf Values

When examining the summary tables, you may encounter special values:

**NaN (Not a Number)** appears when dividing by zero, specifically when:

- Calculating percentage metrics (e.g., `a_markets_flagged_pct` or `na_markets_flagged_pct`) for market types with **zero total markets**
- Example: SE freezers has 0 a_markets, resulting in $\frac{0}{0} \times 100 = \text{NaN}$
- **Interpretation**: "This market type category doesn't exist in this region-product combination"

**-Inf (Negative Infinity)** appears when calculating max() or mean() on an empty set, specifically when:

- Computing excess measures (`max_excess_amount`, `avg_excess_amount`, `avg_excess_pct`) for groups with **zero flagged markets**
- Example: GB freezers a_markets has 0 flagged markets, so `max(excess_amount[flagged])` evaluates to `-Inf`
- **Interpretation**: "No flagged markets exist in this group to calculate statistics from"

These are mathematically correct results, not errors. They indicate categories with insufficient data and should be interpreted as missing or non-applicable values in the analysis context.

**Reporting Options**: You can leave these values as-is for mathematical completeness and transparency about missing data, or replace them with `NA` or `0` depending on your reporting preference and audience needs.

```{r view-flagged-results}
library(ggplot2)
library(knitr)

# Display summary statistics table
cat("\n### Summary Statistics by Region, Product, and Market Type\n\n")
kable(summary_stats, digits = 2, caption = "Flagged Markets Summary Statistics")

# Export summary statistics as formatted table
write_csv(summary_stats, "assets/table1_summary_statistics.csv")

# Display visual summary
cat("\n### Visual Summary by Region and Product\n\n")
kable(visual_summary, digits = 2, caption = "Markets Overview")

# Export visual summary as formatted table
write_csv(visual_summary, "assets/table2_visual_summary.csv")

# Display flagged markets list
cat("\n### List of Flagged Markets (", nrow(flagged_markets), " markets)\n\n")
kable(head(flagged_markets, 10), digits = 2, caption = "Top 10 Flagged Markets by Excess %")

# Export top flagged markets table
write_csv(head(flagged_markets, 10), "assets/table3_top_flagged_markets.csv")

# Plot 1: Flagged markets by region and product
p1 <- ggplot(summary_stats, aes(x = paste(region, product), y = flagged_pct, fill = market_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percentage of Flagged Markets by Region, Product, and Market Type",
       x = "Region & Product",
       y = "Flagged %",
       fill = "Market Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
print(p1)
ggsave("assets/figure1_flagged_pct_by_region_product.png", plot = p1, width = 10, height = 6, dpi = 300)

# Plot 2: Distribution of excess percentages for flagged markets
if (nrow(flagged_markets) > 0) {
  p2 <- ggplot(flagged_markets, aes(x = paste(region, product), y = excess_pct, color = market_type)) +
    geom_point(size = 3) +
    labs(title = "Excess % Above Upper Bound for Flagged Markets",
         x = "Region & Product",
         y = "Excess % Above Upper Bound",
         color = "Market Type") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  print(p2)
  ggsave("assets/figure2_excess_pct_flagged_markets.png", plot = p2, width = 10, height = 6, dpi = 300)
}

# Plot 3: Normalized TCL vs Upper Bound for all markets
p3 <- ggplot(combined_results, aes(x = tcl_upper, y = normalized_tcl, color = flagged)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Normalized Total Consumer Loss vs Upper Bound",
       x = "Upper Bound (tcl_upper)",
       y = "Normalized TCL",
       color = "Flagged") +
  facet_wrap(~ paste(region, product), scales = "free") +
  theme_minimal()
print(p3)
ggsave("assets/figure3_normalized_tcl_vs_upper_bound.png", plot = p3, width = 12, height = 8, dpi = 300)

# Plot 4: Distribution by market size
p4 <- ggplot(combined_results, aes(x = as.factor(mark_size), fill = flagged)) +
  geom_bar(position = "fill") +
  labs(title = "Proportion of Flagged Markets by Market Size",
       x = "Market Size",
       y = "Proportion",
       fill = "Flagged") +
  facet_wrap(~ paste(region, product)) +
  theme_minimal()
print(p4)
ggsave("assets/figure4_proportion_flagged_by_market_size.png", plot = p4, width = 12, height = 8, dpi = 300)
```




