---
title: "EDA Collusion 6"
author: "Ross May"
date: "2026-01-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(123)
```

## Flag homogeneous markets

```{r flag-anomalous-markets}
# Flag markets where normalized total consumer loss exceeds upper bounds
library(dplyr)
library(tidyr)
library(readr)

# Define the configurations for each case
cases <- data.frame(
  region = c("gb", "gb", "se", "se"),
  product = c("freezers", "headphones", "freezers", "headphones"),
  cl_folder = c("cl_results_gb_freezers", "cl_results_gb_headphones", 
                "cl_results_se_freezers", "cl_results_se_headphones"),
  upper_folder = c("upper_pi_bounds_gb_freezers", "upper_pi_bounds_gb_headphones",
                   "upper_pi_bounds_se_freezers", "upper_pi_bounds_se_headphones"),
  upper_file = c("tcl_upper_gb_freez.csv", "tcl_upper_gb_hp.csv",
                 "tcl_upper_se_freez.csv", "tcl_upper_se_hp.csv"),
  stringsAsFactors = FALSE
)

# Function to process a single market file (a_markets or na_markets)
process_market_file <- function(file_path, upper_bounds, region, product, market_type) {
  # Read the market data
  market_data <- read_csv(file_path, show_col_types = FALSE)
  
  # Check if file is empty or has no data rows
  if (nrow(market_data) == 0) {
    return(data.frame())
  }
  
  # Normalize tot_consum_loss by mean of tot_consum_loss
  market_data <- market_data %>%
    mutate(
      normalized_tcl = tot_consum_loss / mean(tot_consum_loss, na.rm = TRUE),
      region = region,
      product = product,
      market_type = market_type
    )
  
  # Merge with upper bounds
  market_data <- market_data %>%
    left_join(upper_bounds, by = c("mark_size" = "ms"))
  
  # Flag markets where normalized TCL exceeds upper bound
  market_data <- market_data %>%
    mutate(
      flagged = normalized_tcl > tcl_upper,
      excess_amount = normalized_tcl - tcl_upper,
      excess_pct = ((normalized_tcl - tcl_upper) / tcl_upper) * 100
    )
  
  return(market_data)
}

# Initialize list to store all results
all_results <- list()

# Process each case
for (i in 1:nrow(cases)) {
  region <- cases$region[i]
  product <- cases$product[i]
  cl_folder <- cases$cl_folder[i]
  upper_folder <- cases$upper_folder[i]
  upper_file <- cases$upper_file[i]
  
  cat(sprintf("\n=== Processing: %s %s ===\n", toupper(region), product))
  
  # Read upper bounds
  upper_bounds_path <- file.path("output", upper_folder, upper_file)
  upper_bounds <- read_csv(upper_bounds_path, show_col_types = FALSE)
  
  # Process a_markets
  a_markets_path <- file.path("output", cl_folder, "a_markets.csv")
  if (file.exists(a_markets_path)) {
    a_results <- process_market_file(a_markets_path, upper_bounds, region, product, "a_markets")
    
    if (nrow(a_results) > 0) {
      # Save detailed results for a_markets
      output_file <- file.path("output", cl_folder, "a_markets_flagged.csv")
      write_csv(a_results, output_file)
      cat(sprintf("  A-markets: %d markets, %d flagged (%.1f%%)\n", 
                  nrow(a_results), sum(a_results$flagged), 
                  100 * sum(a_results$flagged) / nrow(a_results)))
      
      all_results[[paste0(region, "_", product, "_a")]] <- a_results
    }
  }
  
  # Process na_markets
  na_markets_path <- file.path("output", cl_folder, "na_markets.csv")
  if (file.exists(na_markets_path)) {
    na_results <- process_market_file(na_markets_path, upper_bounds, region, product, "na_markets")
    
    if (nrow(na_results) > 0) {
      # Save detailed results for na_markets
      output_file <- file.path("output", cl_folder, "na_markets_flagged.csv")
      write_csv(na_results, output_file)
      cat(sprintf("  NA-markets: %d markets, %d flagged (%.1f%%)\n", 
                  nrow(na_results), sum(na_results$flagged), 
                  100 * sum(na_results$flagged) / nrow(na_results)))
      
      all_results[[paste0(region, "_", product, "_na")]] <- na_results
    }
  }
}

# Combine all results
combined_results <- bind_rows(all_results)

# Save combined detailed results
write_csv(combined_results, "output/all_markets_flagged_detailed.csv")
cat(sprintf("\n=== Overall Results ===\n"))
cat(sprintf("Total markets analyzed: %d\n", nrow(combined_results)))
cat(sprintf("Total flagged markets: %d (%.1f%%)\n", 
            sum(combined_results$flagged), 
            100 * sum(combined_results$flagged) / nrow(combined_results)))

# Create summary statistics
summary_stats <- combined_results %>%
  group_by(region, product, market_type) %>%
  summarise(
    total_markets = n(),
    flagged_markets = sum(flagged),
    flagged_pct = 100 * sum(flagged) / n(),
    avg_normalized_tcl = mean(normalized_tcl),
    avg_tcl_upper = mean(tcl_upper, na.rm = TRUE),
    max_excess_amount = max(excess_amount[flagged], na.rm = TRUE),
    avg_excess_amount = mean(excess_amount[flagged], na.rm = TRUE),
    avg_excess_pct = mean(excess_pct[flagged], na.rm = TRUE),
    .groups = "drop"
  )

# Save summary statistics
write_csv(summary_stats, "output/flagged_markets_summary.csv")
write_csv(summary_stats, "assets/flagged_markets_summary.csv")
print(summary_stats)

# Create a list of flagged markets for further analysis
flagged_markets <- combined_results %>%
  filter(flagged) %>%
  select(region, product, market_type, product_id, mark_size, 
         tot_consum_loss, normalized_tcl, tcl_upper, 
         excess_amount, excess_pct) %>%
  arrange(region, product, market_type, desc(excess_pct))

# Save flagged markets list
write_csv(flagged_markets, "output/flagged_markets_for_analysis.csv")
write_csv(flagged_markets, "assets/flagged_markets_for_analysis.csv")
cat(sprintf("\nFlagged markets saved to: output/flagged_markets_for_analysis.csv\n"))
cat(sprintf("Number of flagged markets for pricing pattern analysis: %d\n", nrow(flagged_markets)))

# Create market size distribution of flagged markets
ms_distribution <- flagged_markets %>%
  group_by(region, product, market_type, mark_size) %>%
  summarise(
    n_flagged = n(),
    avg_excess_pct = mean(excess_pct),
    .groups = "drop"
  ) %>%
  arrange(region, product, market_type, mark_size)

write_csv(ms_distribution, "output/flagged_markets_by_market_size.csv")
write_csv(ms_distribution, "assets/flagged_markets_by_market_size.csv")

# Create visual summary table
visual_summary <- combined_results %>%
  group_by(region, product) %>%
  summarise(
    a_markets_total = sum(market_type == "a_markets"),
    a_markets_flagged = sum(market_type == "a_markets" & flagged),
    a_markets_flagged_pct = 100 * sum(market_type == "a_markets" & flagged) / sum(market_type == "a_markets"),
    na_markets_total = sum(market_type == "na_markets"),
    na_markets_flagged = sum(market_type == "na_markets" & flagged),
    na_markets_flagged_pct = 100 * sum(market_type == "na_markets" & flagged) / sum(market_type == "na_markets"),
    .groups = "drop"
  )

write_csv(visual_summary, "output/flagged_markets_visual_summary.csv")
write_csv(visual_summary, "assets/flagged_markets_visual_summary.csv")
print(visual_summary)

cat("\n=== Analysis Complete ===\n")
cat("Output files created:\n")
cat("  1. output/all_markets_flagged_detailed.csv - All markets with flagging status\n")
cat("  2. output/flagged_markets_summary.csv - Summary statistics by region/product/type\n")
cat("  3. output/flagged_markets_for_analysis.csv - List of flagged markets for pricing analysis\n")
cat("  4. output/flagged_markets_by_market_size.csv - Distribution by market size\n")
cat("  5. output/flagged_markets_visual_summary.csv - Quick visual summary\n")
cat("  6. Individual *_flagged.csv files in each cl_results_* folder\n")
```

## Results Summary

The analysis identified **8 flagged markets (4.2%)** out of 192 total markets where normalized total consumer loss exceeds the upper prediction interval bounds:

**Breakdown by Region and Product:**
- **GB Freezers**: 0 flagged markets (0% of 17 markets)
- **GB Headphones**: 3 flagged markets (5.1% of 59 markets)
  - 2 in a_markets (5.4% of 37)
  - 1 in na_markets (4.5% of 22)
- **SE Freezers**: 1 flagged market (12.5% of 8 na_markets)
- **SE Headphones**: 4 flagged markets (3.7% of 108 markets)
  - 2 in a_markets (5.0% of 40)
  - 2 in na_markets (2.9% of 68)

These flagged markets exhibit anomalously high consumer losses relative to their upper bounds and are candidates for further pricing pattern analysis to identify potential coordination mechanisms.

## Methodology: Summary Statistics Computations

### 1. Normalization
For each market within a region-product-type combination (e.g., all GB headphones a_markets), the total consumer loss is normalized by dividing by the mean:

$$\text{normalized\_tcl} = \frac{\text{tot\_consum\_loss}}{\text{mean(tot\_consum\_loss)}}$$

This produces a dimensionless measure where values around 1.0 represent markets with average consumer loss, while values >1 indicate above-average loss.

### 2. Flagging Criterion
Markets are flagged when their normalized total consumer loss exceeds the upper prediction interval bound for their market size:

$$\text{flagged} = \begin{cases} 
\text{TRUE} & \text{if normalized\_tcl} > \text{tcl\_upper} \\
\text{FALSE} & \text{otherwise}
\end{cases}$$

### 3. Excess Measures
For flagged markets, we compute two excess measures:

- **Excess Amount**: $\text{excess\_amount} = \text{normalized\_tcl} - \text{tcl\_upper}$
- **Excess Percentage**: $\text{excess\_pct} = \frac{\text{normalized\_tcl} - \text{tcl\_upper}}{\text{tcl\_upper}} \times 100$

### 4. Summary Statistics by Region-Product-Type
The `summary_stats` table groups markets by region, product, and market_type, computing:

- **total_markets**: Count of markets in the group: $n()$
- **flagged_markets**: Count of flagged markets: $\sum(\text{flagged})$
- **flagged_pct**: Percentage flagged: $\frac{\sum(\text{flagged})}{n()} \times 100$
- **avg_normalized_tcl**: Mean normalized TCL: $\text{mean(normalized\_tcl)}$
- **avg_tcl_upper**: Mean upper bound: $\text{mean(tcl\_upper)}$
- **max_excess_amount**: Maximum excess among flagged markets: $\max(\text{excess\_amount}|\text{flagged})$
- **avg_excess_amount**: Mean excess among flagged markets: $\text{mean(excess\_amount}|\text{flagged})$
- **avg_excess_pct**: Mean excess % among flagged markets: $\text{mean(excess\_pct}|\text{flagged})$

### 5. Visual Summary by Region-Product
The `visual_summary` table groups by region and product only, computing:

- **a_markets_total**: Count of a_markets: $\sum(\text{market\_type} == \text{"a\_markets"})$
- **a_markets_flagged**: Count of flagged a_markets: $\sum(\text{market\_type} == \text{"a\_markets"} \land \text{flagged})$
- **a_markets_flagged_pct**: Percentage of a_markets flagged: $\frac{\sum(\text{market\_type} == \text{"a\_markets"} \land \text{flagged})}{\sum(\text{market\_type} == \text{"a\_markets"})} \times 100$
- **na_markets_total**: Count of na_markets: $\sum(\text{market\_type} == \text{"na\_markets"})$
- **na_markets_flagged**: Count of flagged na_markets: $\sum(\text{market\_type} == \text{"na\_markets"} \land \text{flagged})$
- **na_markets_flagged_pct**: Percentage of na_markets flagged: $\frac{\sum(\text{market\_type} == \text{"na\_markets"} \land \text{flagged})}{\sum(\text{market\_type} == \text{"na\_markets"})} \times 100$

### 6. Market Size Distribution
The `ms_distribution` table groups flagged markets by region, product, market_type, and mark_size, showing:

- **n_flagged**: Count of flagged markets at each market size
- **avg_excess_pct**: Average excess percentage at each market size

### Interpreting NaN and -Inf Values

When examining the summary tables, you may encounter special values:

**NaN (Not a Number)** appears when dividing by zero, specifically when:

- Calculating percentage metrics (e.g., `a_markets_flagged_pct` or `na_markets_flagged_pct`) for market types with **zero total markets**
- Example: SE freezers has 0 a_markets, resulting in $\frac{0}{0} \times 100 = \text{NaN}$
- **Interpretation**: "This market type category doesn't exist in this region-product combination"

**-Inf (Negative Infinity)** appears when calculating max() or mean() on an empty set, specifically when:

- Computing excess measures (`max_excess_amount`, `avg_excess_amount`, `avg_excess_pct`) for groups with **zero flagged markets**
- Example: GB freezers a_markets has 0 flagged markets, so `max(excess_amount[flagged])` evaluates to `-Inf`
- **Interpretation**: "No flagged markets exist in this group to calculate statistics from"

These are mathematically correct results, not errors. They indicate categories with insufficient data and should be interpreted as missing or non-applicable values in the analysis context.

**Reporting Options**: You can leave these values as-is for mathematical completeness and transparency about missing data, or replace them with `NA` or `0` depending on your reporting preference and audience needs.

```{r view-flagged-results}
library(ggplot2)
library(knitr)

# Display summary statistics table
cat("\n### Summary Statistics by Region, Product, and Market Type\n\n")
kable(summary_stats, digits = 2, caption = "Flagged Markets Summary Statistics")

# Export summary statistics as formatted table
write_csv(summary_stats, "assets/table1_summary_statistics.csv")

# Display visual summary
cat("\n### Visual Summary by Region and Product\n\n")
kable(visual_summary, digits = 2, caption = "Markets Overview")

# Export visual summary as formatted table
write_csv(visual_summary, "assets/table2_visual_summary.csv")

# Display flagged markets list
cat("\n### List of Flagged Markets (", nrow(flagged_markets), " markets)\n\n")
kable(head(flagged_markets, 10), digits = 2, caption = "Top 10 Flagged Markets by Excess %")

# Export top flagged markets table
write_csv(head(flagged_markets, 10), "assets/table3_top_flagged_markets.csv")

# Plot 1: Flagged markets by region and product
p1 <- ggplot(summary_stats, aes(x = paste(region, product), y = flagged_pct, fill = market_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percentage of Flagged Markets by Region, Product, and Market Type",
       x = "Region & Product",
       y = "Flagged %",
       fill = "Market Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
print(p1)
ggsave("assets/figure1_flagged_pct_by_region_product.png", plot = p1, width = 10, height = 6, dpi = 300)

# Plot 2: Distribution of excess percentages for flagged markets
if (nrow(flagged_markets) > 0) {
  p2 <- ggplot(flagged_markets, aes(x = paste(region, product), y = excess_pct, color = market_type)) +
    geom_point(size = 3) +
    labs(title = "Excess % Above Upper Bound for Flagged Markets",
         x = "Region & Product",
         y = "Excess % Above Upper Bound",
         color = "Market Type") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  print(p2)
  ggsave("assets/figure2_excess_pct_flagged_markets.png", plot = p2, width = 10, height = 6, dpi = 300)
}

# Plot 3: Normalized TCL vs Upper Bound for all markets
p3 <- ggplot(combined_results, aes(x = tcl_upper, y = normalized_tcl, color = flagged)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Normalized Total Consumer Loss vs Upper Bound",
       x = "Upper Bound (tcl_upper)",
       y = "Normalized TCL",
       color = "Flagged") +
  facet_wrap(~ paste(region, product), scales = "free") +
  theme_minimal()
print(p3)
ggsave("assets/figure3_normalized_tcl_vs_upper_bound.png", plot = p3, width = 12, height = 8, dpi = 300)

# Plot 4: Distribution by market size
p4 <- ggplot(combined_results, aes(x = as.factor(mark_size), fill = flagged)) +
  geom_bar(position = "fill") +
  labs(title = "Proportion of Flagged Markets by Market Size",
       x = "Market Size",
       y = "Proportion",
       fill = "Flagged") +
  facet_wrap(~ paste(region, product)) +
  theme_minimal()
print(p4)
ggsave("assets/figure4_proportion_flagged_by_market_size.png", plot = p4, width = 12, height = 8, dpi = 300)
```




